---
title: "ESM 206 Assignment 4"
author: "Cameryn Brock and Allie Hacker"
date: "11/6/2019"
output: html_document
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, message=FALSE, warning=FALSE}

# Attach packages

library(tidyverse)
library(janitor)

# Read in & clean up data

lobster_data <- read_csv("lobster_abundance_sbc_lter.csv", 
                         na = "-99999") %>% 
  clean_names()

# Tidy data by uncounting counts

lobster_tidy <- lobster_data %>% 
  uncount(count)

# Add MPA column to data
lobster_mpa <- lobster_tidy %>% 
  mutate(mpa= ifelse(site %in% c("IVEE", "NAPL"), "MPA", "non-MPA" ))

```
#### Introduction

To be written

#### Data and Methods

To be written

#### Results

Data on California Spiny Lobster abundance was collected at five LTER sites. Two of the sites, Isla Vista and Naples, were established as Marine Protected Areas (MPAs) in 2012. Lobster abundance increased from 2012 to 2018 at all sampling locations, but the two sites in MPAs had the greatest percent increase in lobster abundance (Lobster abundance in Isla Vista and Naples increased 36-fold and 49-fold, respectively).

```{r Part A: Changes in lobster abundance}
# Part A: Changes in annual lobster abundance by site

#get annual count of lobsters for each site
lobster_abundance <- lobster_tidy %>% 
  count(site, year) %>% 
  mutate(mpa= ifelse(site %in% c("IVEE", "NAPL"), "MPA", "non-MPA" ))

# graph lobster abundance

ggplot(lobster_abundance, 
       aes(x = year, 
           y = n, 
           group = site, 
           color = mpa, 
           size = mpa)
       )+
  geom_line()+
  scale_color_manual(values = c("aquamarine3", "indianred3"))+
  scale_size_manual(values = c(1.2, 0.5))+
  theme_minimal()+
  labs(x= "Year", y = "Lobsters abundance")+
  scale_x_continuous(breaks = seq(2012, 2018, 1), expand = c(0,0))+
  scale_y_continuous(breaks = seq(0, 1010, 200), expand = c(0,100))+
  coord_cartesian(clip = 'off')+
  theme(legend.title = element_blank(),
        legend.background = element_rect(color = "black"),
        legend.position = c(0.2, 0.8),
        plot.margin = unit(c(0.5,2.5,0,0),"cm"))+
  annotate("text", 
           x = 2018.05, 
           y = 946, 
           label = "Isla Vista", 
           size = 3, 
           hjust = 0)+
  annotate("text", 
           x = 2018.05, 
           y = 290, 
           label = "Naples", 
           size = 3, 
           hjust = 0)+
  annotate("text", 
           x = 2018.05, 
           y = 164, 
           label = "Mohawk", 
           size = 3, 
           hjust = 0)+
  annotate("text", 
           x = 2018.05, 
           y = 355, 
           label = "Carpenteria", 
           size = 3, 
           hjust = 0)+
  annotate("text", 
           x = 2018.05, 
           y = 54, 
           label = "Arroyo Quemado", 
           size = 3, 
           hjust = 0)

```

**Figure 1.** California Spiny Lobster abundance in the Santa Barbara Channel from 2012 to 2018. Sites shown in tuorquoise were established as Marine Protected Areas (MPAs) in 2012.

```{r}
# Mean Lobster size at MPA vs. non-MPA

lobster_size_mpa <- lobster_mpa %>% 
  group_by(mpa, year) %>% 
  summarise(mean_size_mm = mean(size_mm, na.rm = TRUE),
            sd_size = sd(size_mm, na.rm = TRUE),
            lobster_number = n())

# t-test for 2012 lobster size MPA vs non-MPA

mpa_2012_sample <- lobster_mpa %>% 
  filter(mpa == "MPA", year == "2012") %>% 
  pull(size_mm)

non_mpa_2012_sample <- lobster_mpa %>% 
  filter(mpa == "non-MPA", year == "2012") %>% 
  pull(size_mm)

mpa_2012_ttest <- t.test(mpa_2012_sample, non_mpa_2012_sample)

mpa_2012_ttest

# t-test for 2018 lobster size MPA vs non-MPA

mpa_2018_sample <- lobster_mpa %>% 
  filter(mpa == "MPA", year == "2018") %>% 
  pull(size_mm)

non_mpa_2018_sample <- lobster_mpa %>% 
  filter(mpa == "non-MPA", year == "2018") %>% 
  pull(size_mm)

mpa_2018_ttest <- t.test(mpa_2018_sample, non_mpa_2018_sample)

mpa_2018_ttest

```
Mean lobster size in Marine Protected Areas ( mean, sd) and non-Marine Protected Areas (mean, sd) in 2012 differed significantly (p-value =`r mpa_2012_ttest$p.value`)

```{r}
# Change in size at MPA sites from 2012 to 2018

mpa_size_2012_sample <- lobster_mpa %>% 
  filter(mpa == "MPA", year == "2012") %>% 
  pull(size_mm)

mpa_size_2018_sample <- lobster_mpa %>% 
  filter(mpa == "MPA", year == "2018") %>% 
  pull(size_mm)

mpa_size_ttest <- t.test(mpa_size_2012_sample, mpa_size_2018_sample)

mpa_size_ttest


# Change in size at non-MPA sizes from 2012 to 2018

non_mpa_size_2012_sample <- lobster_mpa %>% 
  filter(mpa == "non-MPA", year == "2012") %>% 
  pull(size_mm)

non_mpa_size_2018_sample <- lobster_mpa %>% 
  filter(mpa == "non-MPA", year == "2018") %>% 
  pull(size_mm)

non_mpa_size_ttest <- t.test(non_mpa_size_2012_sample, non_mpa_size_2018_sample)

non_mpa_size_ttest

```

